---
title: "Exploratory stats"
format: html
editor: visual
---

```{r}

library(pacman)

if (!require("pacman")) install.packages("pacman")

pacman::p_load(
  cowplot,
  xtable,
  renv,
  mice,
  weights,
  gridExtra,
  readxl,
  ggplot2,
  ipumsr,
  dplyr,
  data.table,
  sf,
  raster,
  RSQLite,
  lfe,
  car,
  countrycode,
  patchwork,
  fastDummies,
  stargazer,
  sandwich,
  lmtest,
  haven,
  ggthemes,
  RColorBrewer,
  purrr,
  viridis,
  tidyr,
  data.table,
  here,
  plm,
  spdep,
  robust,
  lfe, 
  doParallel,
  foreach,
  pryr,
  fixest,
  broom,
  multiwayvcov
)

here::i_am("code/explore/summary_stats.qmd")


```

```{r}

panel_univ <- readRDS(here("data", "tmp","panel_univ.RDS"))

```

## Exploratory Statistics

### 1) HOUSE OWNERSHIP & INCOME

#### House ownership along ISEE

```{r}

breaks_vec <- c(0, 10000, 15000, 20000, 25000, 30000, 35000, 40000, 60000, Inf)
labels_vec <- c(
  "0–10 k", 
  "10–15 k", 
  "15–20 k", 
  "20–25 k", 
  "25–30 k", 
  "30–35 k", 
  "35–40 k", 
  "40–60 k", 
  "> 60 k"
)

panel_univ <- panel_univ %>%
  mutate(
  #bucket hhs in bins
    isee_bin = cut(
      isee_value,
      breaks = breaks_vec,
      labels = labels_vec,
      include.lowest = TRUE,
      right = FALSE
    )
  )


sum_panel<- panel_univ %>%
  group_by(isee_bin) %>%
  summarise(
    prop_primary   = mean(houseown1, na.rm = TRUE),
    prop_secondary = mean(houseown2, na.rm = TRUE),
    n_obs          = n(),
    .groups        = "drop"
  ) %>%
  filter(n_obs > 0)


plot_panel <- sum_panel %>%
  dplyr::select(isee_bin, prop_primary, prop_secondary) %>%
  pivot_longer(
    cols      = c("prop_primary", "prop_secondary"),
    names_to  = "Type",
    values_to = "Proportion"
  ) %>%
  filter(!is.na(isee_bin))%>%
  mutate(
    typ = Type , 
    typ = ifelse(typ == "prop_primary", "Owns Primary Residence","Owns more than one house")
)





p_house <- ggplot(plot_panel, aes(
  x     = isee_bin,
  y     = Proportion,
  color = typ,
  group = typ
)) +
  geom_line(size = 1) +
  geom_point(size = 2.5) +
  scale_y_continuous(
    name   = "Share of Households",
    labels = percent_format(accuracy = 1),
    limits = c(0, max(plot_panel$Proportion) * 1.05)
  ) +
  scale_color_manual(
    name   = "",
    values = c(
      "Owns Primary Residence" = "#1B9E77",  # green
      "Owns more than one house"       = "#D95F02"   # orange
    )
  ) +
  labs(
    x      = "ISEE Bracket (EUR)",
    title  = "Primary vs. Secondary Home Ownership by ISEE Bracket",
    subtitle = "Proportion of households in that ISEE range primary vs. own a second home"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.title.x     = element_text(size = 12, face = "bold"),
    axis.text.x      = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.y     = element_text(size = 12, face = "bold"),
    plot.title       = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle    = element_text(size = 11, hjust = 0.5),
    legend.position  = "bottom",
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank()
  )


p_house

ggsave(here("output","isee_homeownership_bins_2016_uni.pdf"), width = 8, height = 5, prop_house16)


```

##### Same but using ner income instead of ISEE as running variable 

```{r}

breaks_inc <- c(0, 10000, 15000, 20000, 25000, 30000, 35000, 40000, 60000, Inf)
labels_inc <- c(
  "0–10 k", 
  "10–15 k", 
  "15–20 k", 
  "20–25 k", 
  "25–30 k", 
  "30–35 k", 
  "35–40 k", 
  "40–60 k", 
  "> 60 k"
)


panel_univ <- panel_univ %>%
  mutate(
    income_bin = cut(
      hh_income_net,
      breaks        = breaks_inc,
      labels        = labels_inc,
      include.lowest= TRUE,
      right         = FALSE
    )
  )


sum_inc <- panel_univ %>%
  group_by(income_bin) %>%
  summarise(
    prop_primary   = mean(houseown1, na.rm = TRUE),
    prop_secondary = mean(houseown2, na.rm = TRUE),
    n_obs          = n(),
    .groups        = "drop"
  ) %>%
  filter(!is.na(income_bin) & n_obs > 0)


plot_inc <- sum_inc %>%
  dplyr::select(income_bin, prop_primary, prop_secondary) %>%
  pivot_longer(
    cols      = c(prop_primary, prop_secondary),
    names_to  = "Type",
    values_to = "Proportion"
  ) %>%
  mutate(
    Type = dplyr::recode(
      Type,
      prop_primary   = "Owns Primary Residence",
      prop_secondary = "Owns Secondary Residence"
    )
  )


p_income <- ggplot(plot_inc, aes(
    x     = income_bin,
    y     = Proportion,
    color = Type,
    group = Type
  )) +
  geom_line(size = 1) +
  geom_point(size = 2.5) +
  scale_y_continuous(
    name   = "Share of Households",
    labels = percent_format(accuracy = 1),
    limits = c(0, max(plot_inc$Proportion) * 1.05)
  ) +
  scale_color_manual(
    name   = "",
    values = c(
      "Owns Primary Residence"   = "#1B9E77",  # green
      "Owns Secondary Residence" = "#D95F02"   # orange
    )
  ) +
  labs(
    x        = "Household Net Income Bracket (EUR)",
    title    = "Home Ownership by Income Bracket",
    subtitle = "Proportion of households owning primary vs. secondary residence"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.title.x     = element_text(size = 12, face = "bold"),
    axis.text.x      = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.y     = element_text(size = 12, face = "bold"),
    plot.title       = element_text(face = "bold", size = 16, hjust = 0.5),
    plot.subtitle    = element_text(size = 11, hjust = 0.5),
    legend.position  = "bottom",
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank()
  )





print(p_income)
ggsave(
  here("output","income_homeownership_bins.pdf"),
  plot   = p_income,
  width  = 8,
  height = 5
)

```

#### House ownership by type of acquisition

```{r}

#table(panel_univ$poss.x)

panel_univ$poss <- ifelse(panel_univ$poss.x == 98 , NA, panel_univ$poss.x)


df_q <- panel_univ %>%
  filter(godabit == 1) %>%
  mutate(
    income_q = ntile(y, 5), 
    income_q = factor(income_q,
                      levels = 1:5,
                      labels = c("Q1\n(Lowest)",
                                 "Q2",
                                 "Q3",
                                 "Q4",
                                 "Q5\n(Highest)"))
  )

df_q <- df_q %>%
  mutate(
    poss_cat = case_when(
      poss == 1 ~ "Bought private",
      poss == 2 ~ "Bought company",
      poss == 3 ~ "Bought public",
      poss == 4    ~ "Inherited",
      poss == 5    ~ "Part Bought / Part Inherited",
      poss == 6    ~ "Gift",
      poss == 7    ~ "Built",
      poss == 8    ~ "Other",
      TRUE          ~ NA_character_
    ),
    # add any missing levels so fct_relevel won't complain
    poss_cat = fct_expand(poss_cat,
                          "Bought private", "Bought company", "Bought public",  "Inherited", "Part Bought / Part Inherited",
                          "Gift", "Built", "Other"),
    poss_cat = fct_relevel(poss_cat,
                           "Bought private", "Bought company", "Bought public", "Inherited",
                           "Part Bought / Part Inherited",
                           "Gift", "Built", "Other")
  )



df_q <- df_q %>%
  filter(!is.na(poss_cat))


shares <- df_q %>%
  group_by(income_q, poss_cat) %>%
  summarise(n = n(), .groups="drop_last") %>%
  mutate(share = n / sum(n)) %>%
  ungroup()


ggplot(shares, aes(x = income_q, y = share, fill = poss_cat)) +
  geom_col(width = 0.75, colour = "white", size = 0.2) +
  scale_y_continuous(labels = scales::percent_format(1)) +
  scale_fill_brewer("Ownership type", palette = "Set2") +
  labs(
    x = "Income quintile",
    y = "Share of households (with godabit = 1)",
    title = "Ownership composition by income quintile",
    subtitle = "Households with godabit == 1; categories of how dwelling was acquired"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    axis.text.x      = element_text(size = 12),
    legend.position  = "bottom"
  )



ggplot(shares, aes(x = income_q, y = share, fill = poss_cat, group = poss_cat)) +
  geom_area(position = "fill", colour = NA) +
  scale_y_continuous(labels = scales::percent_format(1)) +
  scale_fill_brewer("Ownership type", palette = "Set2") +
  labs(
    x = "Income quintile",
    y = "Share",
    title = "Ownership composition by income quintile (area)",
    subtitle = "100% stacked area of acquisition type, godabit == 1"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")






df_inh <- panel_univ %>%
  filter(godabit == 1, poss == 4) %>%        # poss==4 → inherited
  mutate(
    income_q = ntile(y, 5),
    income_q = factor(income_q,
                      levels = 1:5,
                      labels = c("Q1\n(Lowest)",
                                 "Q2", "Q3", "Q4",
                                 "Q5\n(Highest)"))
  )


shares_inh <- panel_univ %>%
  filter(godabit == 1) %>%                   # all relevant households
  mutate(
    income_q = ntile(y, 5),
    income_q = factor(income_q,
                      levels = 1:5,
                      labels = c("Q1\n(Lowest)",
                                 "Q2", "Q3", "Q4",
                                 "Q5\n(Highest)"))
  ) %>%
  group_by(area5, income_q) %>%
  summarise(
    total      = n(),
    inherited  = sum(poss == 4, na.rm = TRUE),
    share_inh  = inherited / total,
    .groups = "drop"
  )


# replace the levels/labels with your real region names
region_labels <- c("North", "East", "South", "West", "Central")

df_inh <- df_inh %>%
  mutate(
    ireg = factor(area5,
                  levels = seq_along(region_labels),
                  labels = region_labels)
  )


ggplot(shares_inh, aes(x = income_q, y = share_inh, fill = factor(area5))) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  scale_y_continuous(labels = scales::percent_format(1)) +
  scale_fill_discrete(
    name   = "Region",
    labels = c("North", "East", "South", "West", "Central")
  ) +
  labs(
    x     = "Income quintile",
    y     = "Share inherited",
    title = "Inherited dwellings by Region & Income Quintile"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom")



################


df_model <- panel_univ %>%
  filter(godabit == 1) %>%
  mutate(
    # Create a 0/1 numeric outcome
    inherited01 = as.numeric(poss == 4),     

    # Income quintiles
    income_q = ntile(y, 5),
    income_q = factor(income_q,
                      levels = 1:5,
                      labels = c("Q1\n(Lowest)",
                                 "Q2", "Q3", "Q4",
                                 "Q5\n(Highest)")),

    # Region factor
    area5 = factor(area5,
                  levels = 1:5,
                  labels = c("North","East","South","West","Central"))
  ) %>%
  dplyr::select(-poss)   


mod_inh <- glm(
  inherited01 ~ income_q * area5,
  data   = df_model,
  family = binomial(link = "logit")
)


preds <- ggpredict(mod_inh, terms = c("income_q","area5"))



ggplot(preds, aes(x = x, y = predicted)) +
  geom_point(size = 3, colour = "#2C3E50") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                width = 0.2, colour = "#2C3E50") +
  facet_wrap(~ group, ncol = 3) +
  scale_x_discrete(name = "Income Quintile") +
  scale_y_continuous(name = "Pr(Inherited)", labels = percent_format(1)) +
  labs(
    title    = "Predicted Probability of Inherited Dwelling\nby Income Quintile & Region",
    subtitle = "Logistic model among godabit == 1 households",
    caption  = "Points = estimate, bars = 95% CI"
  ) +
  theme_bw(base_size = 13) +
  theme(
    strip.background   = element_blank(),
    strip.text         = element_text(face = "bold"),
    panel.grid.minor   = element_blank(),
    panel.grid.major.x = element_blank(),
    plot.title         = element_text(face = "bold", size = 14),
    plot.subtitle      = element_text(size = 11)
  )













################


df_model <- panel_univ %>%
  filter(godabit == 1) %>%
  mutate(
    # binary outcome
    inherited01 = as.integer(poss == 4),
    # income quintile factor
    income_q = ntile(y, 5) %>%
               factor(1:5,
                      labels = c("Q1\n(Lowest)",
                                 "Q2","Q3","Q4",
                                 "Q5\n(Highest)")),
    # area5 as a factor (not a number!)
    area5 = factor(area5,
                   levels = 1:5,
                   labels = c("North","East","South","West","Central"))
  ) %>%
  select(-poss)   # drop the original


mod_inh <- glm(
  inherited01 ~ income_q * area5,
  data   = df_model,
  family = binomial(link = "logit")
)


preds <- ggpredict(mod_inh,
                   terms = c("income_q",   # all 5 quintiles
                             "area5"))     # all 5 factor levels


table(preds$x, preds$group)

newdata <- expand.grid(
  income_q = levels(df_model$income_q),
  area5    = levels(df_model$area5)
)

newdata$income_q <- factor(newdata$income_q, levels=levels(df_model$income_q))
newdata$area5    <- factor(newdata$area5,    levels=levels(df_model$area5))

lp <- predict(mod_inh, newdata, type = "link", se.fit = TRUE)

heat   <- newdata %>%
  mutate(
    fit       = lp$fit,
    se        = lp$se.fit,
    pred      = plogis(fit),
    conf.low  = plogis(fit - 1.96*se),
    conf.high = plogis(fit + 1.96*se),
    pct       = percent(pred, accuracy = 0.1)
  )


ggplot(heat, aes(x = income_q, y = area5, fill = pred)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = pct), size = 3) +
  scale_fill_distiller(
    name    = "Pr(inherited)",
    palette = "Blues",
    labels  = percent_format(1)
  ) +
  labs(
    title = "Modeled Pr(Inherited) by Income Quintile & Area5",
    x     = "Income Quintile",
    y     = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x     = element_text(face = "bold"),
    axis.text.y     = element_text(face = "bold"),
    panel.grid      = element_blank(),
    legend.position = "right"
  )



```

### 2) UNIVERSITY ENROLLMENT AND EDUCATION

#### Uni enroll and father´s education

```{r}

```

#### Uni enroll and mother´s education

```{r}

```

#### Uni enroll and previous own study

```{r}

```

#### Uni enroll and regional location

```{r}

```

### 3) ISEE and house ownership

#### Weigth of house in ISEE

```{r}

```

### 4) Table of summary key vars

#### Overall

```{r}

library(kableExtra)

panel_summary <- panel_univ %>%
  group_by(year) %>%
  summarise(
    indiv    = n_distinct(nquest,  na.rm = TRUE),  
    sex              = mean(sex,             na.rm = TRUE),
    eta              = mean(eta,             na.rm = TRUE),
    studio           = mean(studio,          na.rm = TRUE),
    hh_univ          = mean(hh_univ,         na.rm = TRUE),
    parent_studio    = mean(parent_studio,   na.rm = TRUE),
    own_house        = mean(own_house,       na.rm = TRUE),
    hh_income_net    = mean(hh_income_net,   na.rm = TRUE),
    valabit          = mean(valabit,         na.rm = TRUE),
    fin_net_assets   = mean(fin_net_assets,  na.rm = TRUE)
  ) %>%
  ungroup()


panel_summary <- panel_summary %>%
  mutate(across(-c(year, indiv), ~ round(.x, 2)))


table <- panel_summary %>%
  rename(
    Year               = year,
    `N`  = indiv,
    Sex                 = sex,
    Age                 = eta,
    Study               = studio,
    `HH Univ`           = hh_univ,
    `Parent Study`      = parent_studio,
     `Owner`             = own_house,
    `HH Income Net`     = hh_income_net,
    `Valabit`           = valabit,
    `Fin Net Assets`    = fin_net_assets
  ) %>%
  kbl(
    format       = "latex",
    booktabs     = TRUE,
    caption      = "Descriptive Statistics by Year",
    label        = "tab:desc_stats"
  ) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),
    font_size     = 8
  ) %>%
  pack_rows("Summary", 1, nrow(panel_summary))

cat(table)
```

### Uni enrollment by income quintile and tenure

```{r}


library(forcats)


quintile_plot_df <- panel_univ %>%
  filter(!is.na(hh_univ), !is.na(hh_income_net), !is.na(own_house)) %>%
  mutate(
    income_quintile = ntile(hh_income_net, 5),
    income_quintile = factor(income_quintile,
                              levels = 1:5,
                              labels = paste0("Q",1:5)),
    # turn the 0/1 variable into a proper factor with labels
    tenure = factor(own_house,
                    levels = c(0, 1),
                    labels = c("Renters", "Owners"))
  ) %>%
  group_by(income_quintile, tenure) %>%
  summarise(
    share_hh_univ = mean(hh_univ, na.rm = TRUE),
    .groups = "drop"
  )


quint <- ggplot(quintile_plot_df, aes(
    x    = income_quintile,
    y    = share_hh_univ,
    fill = tenure
  )) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(aes(label = percent(share_hh_univ, accuracy = 1)),
            position = position_dodge(width = 0.7),
            vjust = -0.3,
            size = 3) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.1))
  ) +
  scale_fill_manual(
    values = c("Renters" = "#1f77b4", "Owners" = "#ff7f0e"),
    name   = ""
  ) +
  labs(
    title    = "University Enroll. by Income Quintile & Tenure",
    subtitle = "Share of households with at least one 18 - 30 year-old in university",
    x        = "Income Quintile (1 = lowest)",
    y        = "Share of Households",
    caption  = "Source: Bank of Italy Survey"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title       = element_text(face = "bold", size = 14, hjust = 0.5),
    plot.subtitle    = element_text(size = 10, hjust = 0.5, margin = margin(b = 10)),
    axis.title       = element_text(face = "bold"),
    axis.text.x      = element_text(vjust = 0.5, hjust = 0.5),
    legend.position  = "bottom",
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_blank()
  )



ggsave(here("output","quintile_uni.pdf"), quint)
```

### 
