---
title: "Exploratory stats"
format: html
editor: visual
---

```{r}

library(pacman)

if (!require("pacman")) install.packages("pacman")

pacman::p_load(
  cowplot,
  xtable,
  renv,
  mice,
  weights,
  gridExtra,
  readxl,
  ggplot2,
  ipumsr,
  dplyr,
  data.table,
  sf,
  raster,
  RSQLite,
  lfe,
  car,
  countrycode,
  patchwork,
  fastDummies,
  stargazer,
  sandwich,
  lmtest,
  haven,
  ggthemes,
  RColorBrewer,
  purrr,
  viridis,
  tidyr,
  data.table,
  here,
  plm,
  spdep,
  robust,
  lfe, 
  doParallel,
  foreach,
  pryr,
  fixest,
  broom,
  multiwayvcov
)

here::i_am("code/explore/summary_stats.qmd")


```

```{r}

panel_univ <- readRDS(here("data", "tmp","panel_univ.RDS"))

```

## Exploratory Statistics 

### 1) HOUSE OWNERSHIP  & INCOME

####  House ownership along Income

```{r}

```

#### House ownership along ISEE

```{r}

```

### 2) UNIVERSITY ENROLLMENT AND EDUCATION

#### Uni enroll and father´s education 

```{r}

```

#### Uni enroll and mother´s education 

```{r}

```

#### Uni enroll and previous own study

```{r}

```

#### Uni enroll and regional location 

```{r}

```

### 3) ISEE and house ownership 

#### Weigth of house in ISEE

```{r}

```

### 4) Table of summary key vars 

#### Overall

```{r}

```

#### Balance observables owners and not owners

```{r}

```
