---
title: "create panel"
format: html
editor: visual
---

```{r}

library(pacman)

if (!require("pacman")) install.packages("pacman")

pacman::p_load(
  cowplot,
  xtable,
  renv,
  weights,
  gridExtra,
  readxl,
  ggplot2,
  ipumsr,
  dplyr,
  data.table,
  sf,
  raster,
  RSQLite,
  lfe,
  car,
  countrycode,
  patchwork,
  fastDummies,
  stargazer,
  sandwich,
  lmtest,
  haven,
  ggthemes,
  RColorBrewer,
  purrr,
  viridis,
  tidyr,
  data.table,
  here,
  plm,
  spdep,
  robust,
  lfe, 
  doParallel,
  foreach,
  pryr,
  fixest,
  broom,
  multiwayvcov
)

here::i_am("code/clean/create_panel.qmd")

```

Using Bank of Italy survey data open samples for all available years from :

https://www.bancaditalia.it/statistiche/tematiche/indagini-famiglie-imprese/bilanci-famiglie/distribuzione-microdati/index.html?page=1

(storico_stata file , updated 10/01/2025)

\-\-- \> Storico does not have UNIVER question so go back to using year by year samples : 2010-2012-2014-2016-2020 (5 waves)

-   here opening storico and realizing no vars

```{r}

## unzip

zip_path <- here("data", "raw", "storico_stata.zip")

utils::unzip(zip_path, list = TRUE)
out_dir <- here("data", "raw", "storico_stata_unzipped")
if (!dir.exists(out_dir)) dir.create(out_dir)

utils::unzip(zip_path, exdir = out_dir)

list.dirs(out_dir, full.names = TRUE, recursive = FALSE)
list.files(out_dir, recursive = TRUE)


## open files in env

stata_dir <- here("data", "raw", "storico_stata_unzipped", "STATA")

dta_files <- list.files(stata_dir, pattern = "\\.dta$", full.names = TRUE)


data_list <- lapply(dta_files, read_dta)
names(data_list) <- tools::file_path_sans_ext(basename(dta_files))

list2env(data_list, envir = .GlobalEnv)

ls()

```

-   open year by year

    ```{r}


    Git reset—hard 2f5fca21a4db6a32f954495c4a082cc8082e5c8e 


    years <- c("2010", "2012", "2014", "2016", "2020")


    data_list <- list()


    for (yr in years) {

      dir_path <- here("data", "raw", yr)
      
      files <- list.files(dir_path, pattern = "\\.dta$", full.names = TRUE)

      for (f in files) {
        obj_name <- paste0(yr, "_", tools::file_path_sans_ext(basename(f)))
        data_list[[obj_name]] <- read_dta(f)
      }
    }


    list2env(data_list, envir = .GlobalEnv)


    ls(pattern = "^(2010|2012|2014|2016|2020)_")



    ```

-   Isolate panel component merging components infos

    ```{r}



    years <- c("2010","2012","2014","2016","2020")
    base_mods <- c("carcom","ricfam","rfam")

    panel_long <- map_df(years, function(yr) {
      yr2 <- substr(yr, 3, 4)
      mods <- c(paste0(base_mods, yr2), paste0("q", yr2, "a"), paste0("q", yr2, "d"), "alld1", "alld2_res")
      dir  <- here("data","raw",yr)
      files <- list.files(dir, "\\.dta$", full.names = TRUE)
      keep <- tools::file_path_sans_ext(basename(files)) %in% mods
      files <- files[keep]
      dfs <- map(files, read_dta)
      names(dfs) <- tools::file_path_sans_ext(basename(files))
      car <- dfs[grep(paste0("^carcom",yr2,"$"), names(dfs))]
      oth <- dfs[names(dfs) != paste0("carcom",yr2)]
      df_year <- reduce(car, full_join, by = c("nquest","nordp"))
      df_year <- reduce(oth, full_join, by = "nquest", .init = df_year)
      df_year %>% mutate(year = as.integer(yr))
    })







    ```

## Clean samples

```{r}


## info house

case_16 <- q16d %>%
  dplyr::select(nquest, anabit, anposs, ancostr, godabit, valabit,poss, altrab, altrfab, teragr, ternagr, debita1, ndebita1, debita2, ndebita2, mutuor)


## info household level

hhs <- comp %>%
  dplyr::select(nquest, nord, ncomp, sesso, anasc, staciv, studio, annoedu, pesofit, tipodip, nascarea, univer, parent, ireg, nascreg, area5, acom5, perl, nperl, eta, isco)

income <- rfam %>%
  dplyr::select(y, yl,ym, yt,yc, nquest)

wealth  <- ricfam %>%
  dplyr::select(w, ar1, ar, af,pf, clw,clw2, bd, nquest)%>%
  mutate(
    fin_assets = ar,
    debts = bd
  ) 


# Merge datasets

fam_16 <- hhs_16 %>%
  left_join(wealth_16, by = "nquest") %>%
  left_join(income_16, by = c("nquest")) %>%
  left_join(case_16, by = "nquest")



########################################################################

```

Compute variables for analysis

-   ISEE proxy

```{r}

```

-   University attendance at hh level

    ```{r}

    ```

-   House ownership

    ```{r}

    ```

-   ISEE without house ownership

    ```{r}

    ```

-   Other controls useful for matching

    ```{r}






    ```

## Save panel as a tmp

```{r}

```
